String automata represent a {\em mechanical} description of languages over
$\alphabet$ (\ie subsets of the strings which may be formed from the
characters of a particular alphabet, $\alphabet$).  There are numerous
equivalent ways of stating this, possibly the most useful being that an
automaton can be thought of as an {\em indicator function} of type $\alphabet^*
\to 2$, signaling whether a given string is or is not part of its language.

\subsection{Execution of Automata}

At the core of each automaton will be a \defn{transition function},
$\delta$, which will take the current configuration of the machine (which we
shall denote in general with $c \in \config$) and a character of $\alphabet$ to
produce a new configuration (usually very closely related to the input
configuration, having been changed through a constrained interface).  We say
that $\delta$ has type $\alphabet \times \config \to \config$, sometimes
written $\delta : \alphabet \times \config \to \config$.  Each automaton will
specify exactly one \defn{initial configuration}, $c_0 \in \config$, and
will designate some configurations as \defn{accepting configurations},
$\config_F \subseteq \config$.

To test whether a particular string $\vec{s} = s_1 \ldots s_n \in \alphabet^*$
is in its language, one starts with the automaton in its initial
configuration, feeds its transition function each character of the string in
turn (\ie $s_1$, then $s_2$, and so on, up to $s_n$), and then looks to see
if the automaton is now in an accepting configuration.  More formally, we
can define a \defn{run} of an automaton as a string composed of alternating
elements of $\config$ and $\alphabet$.  If, after transitioning on $s_i$ the
machine is in configuration $c_i$, then $c_0 s_1 c_1 s_2 c_2 \ldots s_n c_n$
is the run of the machine on $\vec{s} \in \alphabet^*$.  A run is said to be
accepting if $c_n$ is an accepting configuration (\ie $c_n \in \config_F$).%
%
\footnote{It is equally sensible to feed strings through an automaton in
reverse and ask if, after receiving the first symbol of the (un-reversed)
string the machine is in an accepting configuration.}

\subsection{Nondeterministic Automata}

The above scenario described a \defn{deterministic} automaton, which always
make a particular move given a configuration and an input character.
Nondeterministic automata instead may take on {\em one of many}
configurations after receiving a character (as well as at initialization).
To model these, we define the transition function to result in a {\em set}
of configurations.%
%
\footnote{Many texts instead use a \defn{transition relation}.  While
equivalent, we find it useful to distinguish the inputs and outputs of the
transition process.}
%
Now a run of our nondeterministic machine looks like $C_0 s_1 C_1 \ldots s_n
C_n$ where each $C_i \subseteq \config$ and $c' \in C_{i+1}$ iff $\exists_{c
\in C_i} . c' \in \delta\paren{c,s_{i+1}}$.  We consider the string
$\vec{s}$ accepted by the automaton if $C_n \cap \config_F \ne \emptyset$;
equivalently, we may say that a string is accepted if {\em there exists} a
deterministic run, as defined earlier, where each $c_i \in C_i$, and $c_n
\in \config_F$.

\subsection{Transducers}

A useful generalization of automata is from acceptors ($\alphabet^* \to 2$) to
transducers ($\alphabet^* \to \alphabet'^*$).  Here, $\delta$ not only outputs a
new configuration, it additionally outputs (sometimes optionally) a
character of the \defn{output alphabet} $\alphabet'$.  Nondeterministic
transducers' transition functions typically output a set of
configuration/character pairs ($(c, s') \in \config \times \alphabet'$), but
may sometimes be allowed to skip an emission (making $\delta : \alphabet \times
\config \to \config \times (1+\alphabet')$) or to produce multiple output
characters (\eg $\delta : \alphabet \times \config \to \config \times
\alphabet'^*$).

\Note{Explain ...}
\begin{itemize}
  \item Sequential
  \item Unambiguous
\end{itemize}

\subsection{Weighted Automata}

Related to transducers are weighted automata, which may be thought of as
scoring or ranking input strings.  That is, rather than being indicator
functions $\alphabet^* \to 2$, they capture functions from $\alphabet^*$ to a
semi-ring, $R = \paren{0_R, 1_R, +_R, \times_R}$, such as $\mathbb{N}$ or
$\mathbb{R}$.  The typical definition of a weighted string automata has
$\delta$ producing an element of $R$ on each transition and adds two new
functions, which assign weights to the initial ($\config_0 \to R$) and
final ($\config_F \to R$) states.%
%
\footnote{For weighted {\em deterministic} automata, there is only one
initial configuration and so one typically elides the former function.}
%
The weight of a run is then the
$R$-product of its initial and final states' weights and the weights of all
transitions therein.  The weight of a string is the $R$-sum of all accepting
runs for that string.%
%
\footnote{Equivalently, we can think of a typical weighted automaton as a
non-deterministic {\em transducer to $R^*$} followed by per-run $R$-product
and cross-run $R$-sum reductions.}

\subsection{Descriptive Taxonomy}

We have said that an automaton's transition function $\delta$ has type
$\config \times \alphabet \to \config$, but that the input and output
configurations are often closely related.  In such a case, we may say that
$\delta$ is \defn{characterized} by a function of a different return
type.  What we mean by this is that the information bottleneck represented
by this alternate return type is sufficient to describe the change between
elements of $\config$ on a transition.  See the entry for
\hyperref[sec:zoo-str/pda]{Push-down Automata} for a simple example.

\subsection{Operational Taxonomy}

In an effort to neatly summarize many properties of string automata, entries
will typically be a brief prose description of the family followed by a
table containing common properties.  Uncommon properties or note-worthy
features will be left to prose.

Here, we use the common table to describe itself.  Given a string $s \in
\alphabet^*$, automata $A, A', \ldots$ and their languages $\alang{A},
\alang{A'}, \ldots$ ($\subseteq \alphabet^*$)...
\autinfo{
  member={$s \in \alang{A}$?},
  empty={$\alang{A} = \emptyset$?},
  univ={$\alang{A} = \alphabet^*$?},
  finite={$\abs{\alang{A}} < \infty$?},
  equiv={$\alang{A} = \alang{A'}$?},
  subset={$\alang{A} \subseteq \alang{A'}$?},
%
  kstar={Find $B$ s.t. $\alang{B} = \alang{A}^*$},
  kplus={Find $B$ s.t. $\alang{B} = \alang{A}^+$},
  compl={Find $B$ s.t. $\alang{B} = \alphabet^* \setminus \alang{A}$},
  relcompl={Find $B$ s.t. $\alang{B} = \alang{A} \setminus \alang{A'}$},
%
  intersect={Find $B$ s.t. $\alang{B} = \alang{A} \cap \alang{A'}$},
  union={Find $B$ s.t.  $\alang{B} = \alang{A} \cup \alang{A'}$},
%
  hom={Find $B$ for hom. $h$ s.t. $\alang{B} = h\paren{\alang{A}}$},
  invhom={Find $B$ for hom. $h$ s.t. $\alang{A} = h\paren{\alang{B}}$},
  efhom={Find $B$ for $\epsilon$-free hom. $h$ s.t. $\alang{B} =
         h\paren{\alang{A}}$},
%
  reginter={Find $B$ for \hyperref[sec:zoo-str/fsm]{regular language}
            $L$ s.t. $\alang{B} = \alang{A} \cap L$.},
%
  determinize={Find a deterministic $B$ with language equal to
             (nondet.) $A$.},
  minimize={Find a ``smallest'' $B$ with language equal to $A$.%
            %
            \footnote{``Smallest'' is not always straightforward to define.
            The goal, loosely, is to minimize the amount of information
            necessary to describe the automaton's configuration.}},
}

Of course, not every row in the table is independent of the others.  The
following implications (at least) hold in all cases:
\begin{itemize}
	\item Subset testing implies equivalence testing, of which universality
          and emptiness testing are often special cases (some automata
          families may not be able to represent $\emptyset$ or
          $\alphabet^*$).
	\item Relative complement closure implies general complement closure, assuming $\alphabet^*$ is expessible.
	\item Intersection and general complement closures imply
          relative complement closure.
    \item If a class has relative complement closure and a decidable emptiness test,
          then it has a decidable subset test:
          $S \subseteq T \Leftrightarrow S \setminus T = \emptyset$.
	\item Closure under arbitrary homomorphism implies closure under
          $\epsilon$-free homomorphisms.
\end{itemize}
