Reduction Automata were introduced in \cite{dauchet:reduction}.

\Note{RA is single-ply AWEDC subject to Reduction metaconstraint (and, by
implication, Stated).}

\Note{\cite[p133]{tata} discusses the relaxation of RA to permit AWCBB-style
constraints freely; what I've called ``Deep-Reduction'' above.  This class has decidable emptyness testing.}

\subsection{Proof of Decidable Emptiness of Deterministic RA}

\cite[\S 4.1]{dauchet:reduction} offers an interesting proof that emptiness
testing for deterministic RAs is decidable.  \Note{...}

\subsection{Proof of Non-deterministic RA Undecidable Emptiness Testing}
\label{sec:zoo-tree/ra:ndempty}

\cite[Thm 4.4.7]{tata} contains a proof%
%
\footnote{Apparently first presented in \cite[Appendix
C]{jacquemard:tamodeq} and in contradiction to the statement made in
\cite[\S 4.1]{dauchet:reduction} that the proof therein, which applies to
{\em deterministic} and {\em complete} RAs, carries over to the
non-deterministic case as well.}
%
that emptiness is not decidable for the class of nondeterminstic RAs.  (This
proof generalizes to a large class of tree automata; see
\autoref{sec:tree-sepex:twocm}.)  We feel that a few words said about this
proof here may help to de-mystify it.  A decomposition if this machine into
simpler automata whose intersection is this machine may be found in
\ref{sec:tree-sepex:twocm}.

\begin{wrapfigure}{r}{3.5in}\centering\begin{tikzpicture}
\Tree [.h$_n$
         [.g $C_n$
            [.h$_{n-1}$ [.g $C_{n-1}$ [.h$_{n-2}$ \edge[roof];$\cdots$ ] ]
                           [.h$_{n-2}$ \edge[roof];$\cdots$ ] ] ]
         [.h$_{n-1}$
			[.g \edge[roof];$\cdots$ ]
			[.h$_{n-2}$
				[.g \edge[roof];$\cdots$ ]
				[.$\cdots$	$\cdots$ [.h$_{1}$ [.g \edge[roof];$C_1$ \# ] \# ] ]
			]
         ]
      ]
\end{tikzpicture}\end{wrapfigure}

The overall structure of the trees accepted by this construction was not
immediately obvious to at least one author of this zoo, so a schematic view
is given here.  Each h$_i$ represents the apex of {\em the same tree}
everywhere it occurs; the subscript is {\em not} part of the functor symbol,
which is just h at each of these positions.%
%
\footnote{\cite{tata} sometimes labels the root of the tree ``k'' and
sometimes ``h$'$''; this does not appear essential to the construction and
so we simply use ``h'' for the entire right spine; the only requirement
imposed by the Reduction metaconstraint is that root {\em state} have rank
greater than the state label used for the right spine and for the top-left q
node; this does not require that the root {\em symbol} itself be different.}
%
Every h$_i$ node carries the trace of the first $i$ steps of execution on
its right; its left child can be thought of as the pair of the $i$-th state
of the machine's execution, denoted $C_i$, and h$_{i-1}$.  The equality
constraints are introduced only at the top and at ``g'' nodes in the tree;
the state labels are carefully constructed so that only three ranks are
required to satisfy the Reduction metaconstraint: a base rank, which
includes all labels in the $C_i$ trees is $\le$ a ``g''-state rank which
includes the labels of all ``g'' and ``h'' nodes in the tree except the
apex; the accepting state at the root is the sole member of the third rank.
There will be $2^{n-i}$ copies of $h_i$ (or $C_i$) in a tree encoding $n$
steps.  This construction is {\em deeply} dependent on the fact that
non-determinism allows equal trees to have several distinct runs. 
