A push-down automaton (PDA) is a \hyperref[sec:zoo-str/fsm]{FSM} equipped
with a stack.  Let $\Gamma$ be the set of symbols on the stack and $\sts$
the set of states of the FSM, as before; then $\config = \mathcal{Q} \times
\Gamma^*$.  Initial configurations are typically defined, WOLOG, to have an
empty stack.

Each transition of the automaton is given the top symbol of the
stack as well as the FSM's state and input character, and may manipulate the
stack by optionally popping the top symbol, and optionally pushing a new
symbol.  That is, $\delta$ may be characterized by the type $\alphabet
\times \mathcal{Q} \times \paren{1 + \Gamma} \to \mathcal{Q} \times
\paren{1 + \Gamma} \times 2$, where $1 + \Gamma$ indicates an optional push
to the stack and $2$ indicates a boolean decision to pop.%
%
\footnote{Attempting to pop an emtpy stack is assumed to leave the stack
empty.  We might have been more precise by forbidding popping an empty stack
and instead have said that $\delta$ was fully characterized by the type
$\paren{\alphabet \times \sts \to \sts \times \paren{1 + \Gamma}} \times
\paren{\alphabet \times \sts \times \Gamma \to \sts \times \paren{1 + \Gamma}
\times 2}$, but this more-strict view does not seem worth-while.}
%

There are two, equivalent \cite{xxx}, definitions of accepting
configurations for PDAs:
%
\begin{itemize}
%
    \item The ``accepting state'' condition requires that the final
    configuration be in an accepting $\mathcal{Q}_f$, but may have arbitrary stack
    contents.%
    %
    \footnote{That is, $\pi_1 c_n \in \mathcal{Q}_f$.}
%
    \item The ``emtpy stack'' condition requires only that the stack be
    empty at the end of the string, and does not care in which state of
    $\mathcal{Q}$ the machine finds itself.
    %
    \footnote{That is, $\pi_2 c_n = \epsilon$.}
%
\end{itemize}

\autinfo{
    empty={Yes \cite{xxx}},
}

\subsection{Non-deterministic PDAs}
\label{sec:zoo-str/pda-nd}

Non-deterministic PDAs' recognition corresponds exactly to context free languages.
Deterministic PDAs can recognize the subset of context-free languages,
appropriately called ``deterministic context free languages''.
\Note{cites and hrefs}

See \href{http://wikipedia.org/Push-down\_automaton}{Wikipedia's treatment}
or \cite[Ch. 2]{sipser:theorycomp} for thourough introductions.

\autinfo{
    empty={Yes \cite{xxx}},
    univ={Undecidable; proof via ``computation history'' of a
          \hyperref[sec:zoo-str/tm]{TM}. \cite{xxx}},
    equiv={Undecidable; $\alphabet^*$ is recognizable, but universality is
           undecidable.},
    intersect={Undecidable; reduction from Post Correspondance \cite{xxx}},
}

