\paragraph{Turing Machine}
A FSM augmented with an unbounded tape.  A model of arbitrary computation.

Let $\Gamma$ denote the tape alphabet (with $\Sigma \subseteq \Gamma$), then
$\config = \sts \times \mathbb{N} \times \Gamma^\mathbb{N}$, representing
the FSM state, the current position on the tape, and the tape contents.%
%
\footnote{Unlike most of the other automata in the zoo, the input to a TM as
defined here is not given character-by-character, but is encoded into the
seed machine configuration.  Of course, an equivalent formulation, allowing
the machine to await the next character of its input, is possible, but note
that there remains some fundamental level of asynchrony -- the machine may
compute without bound between requests for the next character.}

$\delta$ is characterized by the type $\paren{\sts \times \Gamma} \to
\paren{\sts \times \Gamma \times 3}$; the input $\Gamma$ is the symbol at
the current position, while the output $\Gamma$ will be written there in the
next configuration; the $3$ indicates a choice between incrementing,
decrementing (if nonzero), or not changing the current position.%
%
\footnote{Attempting to decrease while at position $0$ is the same as a
no-operation; again, we could have given a more precise characterization of
$\delta$ with the attendant loss of clarity.  Note, however, that a machine
with a tape indexed by $\mathbb{Z}$ has no such corner case and is equivalent
to the presentation given here. \cite{post:1936}}

\autinfo{
    empty={Undecidable},
    member={Semidecidable},
    intersect={Yes},
    regops={Yes},
    kplus={Yes},
    determinize={Yes, at exponential space and time cost},
    minimize={No; the set of minimal TMs is not even recognizable.}
}
